---
title: "The R Project"
author: "Jocelyn Garay"
date: "2025-08-06"
categories: [projects]
---

This is my first project post using R.

------------------------------------------------------------------------

# Introduction

This study will be focused on the interaction between hours of music spent listening and its impact on multiple mental health categories.

This project will examine the relationship between the amount of time spent listening to music and its impact on various mental health factors, including anxiety, obsessive-compulsive disorder (OCD), depression, and insomnia. Specifically, this study aims to determine whether increased music listening is associated with lower levels of anxiety, OCD symptoms, depressive symptoms, and insomnia.

We hypothesize that higher duration of music listening will correlate with reduced levels of anxiety and OCD. Additionally, we predict that greater time spent listening to music will be associated with lower levels of depression and fewer symptoms of insomnia.

------------------------------------------------------------------------

# Approach

### Data Set

Data collection was administrated by a college undergraduate student, Catherine Rasgaitis, at the University of Washington. Data was collected in various social media platforms, some including posted on Reddit posts and Discord servers. There were over 700 participants that responded to the survey.

### **Tools**

For this project, we used R and RStudio through Posit Cloud. Other website hosting platforms were used, such as GitHub.

### **Approach/Predictions**

I will examine the correlation between hours of listening to music and mental health conditions between anxiety, depression, insomnia, and OCD.

------------------------------------------------------------------------

# Results

#### Question 1: Does the amount of time spend listening to music impact anxiety levels?

```{r}
#| message: false
#| warning: false
library(ggplot2)

mydata <- read.csv("../../Music_x_Mental_Health_Data_Results.csv")

ggplot(data = mydata, 
       mapping = aes(x= Hours.per.day, y = Anxiety, 
  color = Music.effects, shape = Music.effects)) + 
  geom_point(alpha = 0.8) + geom_smooth(se= FALSE) + 
  labs(
  title = "Relationship between Hours of Music Listening and Anxiety",
  x = "Hours Spent Listening to Music (0-25)",
  y = "Self-Reported Anxiety (0-10)"
) + theme_classic()
```

#### Question 2: Does the amount of time spend listening to music impact depression levels?

```{r}
#| message: false
#| warning: false
library(ggplot2)

mydata <- read.csv("../../Music_x_Mental_Health_Data_Results.csv")

ggplot(data = mydata, 
       mapping = aes(x= Hours.per.day, y = Depression, 
  color = Music.effects, shape = Music.effects)) + 
  geom_point(alpha = 0.8) + geom_smooth(se= FALSE) + 
  labs(
  title = "Relationship between Hours of Music Listening and Depresion",
  x = "Hours Spent Listening to Music (0-25)",
  y = "Self-Reported Depression (0-10)"
) + theme_classic()
```

#### Question 3: Does the amount of time spend listening to music impact insomnia levels?

Here's what was found:

```{r}
#| message: false
#| warning: false
library(ggplot2)

mydata <- read.csv("../../Music_x_Mental_Health_Data_Results.csv")

ggplot(data = mydata, 
       mapping = aes(x= Hours.per.day, y = Insomnia, 
  color = Music.effects, shape = Music.effects)) + 
  geom_point(alpha = 0.8) + geom_smooth(se= FALSE) + 
  labs(
  title = "Relationship Between Hours of Music Listening and Insomnia",
  x = "Hours Spent Listening to Music (0-25)",
  y = "Self-Reported Insomnia (0-10)"
) + theme_classic()
```

#### Question 4: Does the amount of time spend listening to music impact OCD levels?

```{r}
#| message: false
#| warning: false
library(ggplot2)

mydata <- read.csv("../../Music_x_Mental_Health_Data_Results.csv")

ggplot(data = mydata, 
       mapping = aes(x= Hours.per.day, y = OCD, 
  color = Music.effects, shape = Music.effects)) + 
  geom_point(alpha = 0.8) + geom_smooth(se= FALSE) + 
  labs(
  title = "Relationship between Hours of Music Listening and OCD",
  x = "Hours Spent Listening to Music (0-25)",
  y = "Self-Reported OCD (0-10)"
) + theme_classic(base_size = 12)
```

------------------------------------------------------------------------

# Discussion
